---
title: "Rating Shift Happens"
subtitle: "Documentation"
author: "Felipe MENDEZ and Benjamin RENARD (INRAE, RiverLy and RECOVER)."
date: "December 2023"
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/readme/README-",
  out.width = "100%"
)
```

## Introduction

The goal of `RatingShiftHappens` package is to create a tools package for detecting, visualizing and estimating rating shifts. This documentation provides the description of several functions available to the segmentation process. 

Three fundamental functions are available so far :


1. segmentation.engine
2. segmentation
3. recursive.segmentation

## Installation

You can install the development version of [RatingShiftHappens](https://github.com/Felipemendezrios/RatingShiftHappens) from GitHub using the following command. Please note that the [RBaM](https://github.com/BaM-tools/RBaM) package is also required to run this package.

``` {r  results='hide',message=FALSE}
# install.packages("devtools")
devtools::install_github("Felipemendezrios/RatingShiftHappens")
devtools::install_github('BaM-tools/RBaM') 

library(RatingShiftHappens)
```

Functions will be explained more precisely below along with a example.

##  Segmentation procedure for a **known** given number of segments

This is a basic example which shows you how to segment a random variable:

```{r segmentation.engine}
 # Set random generation
 set.seed(1)

 # Create observation vector
 obs=c(rnorm(25,mean=0,sd=1),rnorm(25,mean=2,sd=1))
 nS.user=2

 # Run segmentation engine function
 res <- segmentation.engine(obs=obs,nS=nS.user)

 # Estimated shift time
 res$tau

 # intervals defined by time shifts
 if(nS.user!=1){
  intervals.time.shift=c(res$data.p$time[[1]][1],res$tau,rev(res$data.p$time[[nS.user]])[1])
 }else{
  intervals.time.shift=list(res$data.p$time[1],rev(res$data.p$time)[1])
 }

 # Maximum a posterior value per segment indexed by the list number
 res$segments

 # Uncertainty in shift time
 if(nS.user!=1){
    Shift=res$mcmc$tau1
    hist(Shift,
         main='Histogram of first shift')

    uncertainty95_shift <- list()
    for(i in 1:(nS.user-1)){
     uncertainty95_shift[[i]] = stats::quantile(res$mcmc[,nS.user+i],probs=c(0.025,0.975))
    }
 }


 # Uncertainty in segment estimation
 mu.seg.1.unc=res$mcmc$mu1
 graphics::hist(mu.seg.1.unc,
                xlab='obs',
                main='Histogram of first segment of observation')

 uncertainty95_segment <- list()
 for(i in 1:nS.user){
    uncertainty95_segment [[i]] = stats::quantile(res$mcmc[,i],probs=c(0.025,0.975))
 }

 # Separate and assign information by identified stable period
 res$data.p
 
 # DIC estimation
 res$DIC

 # Setting transparency for plotting
 alpha = 125

 # Set color plot
 color_customized_rect <- function(alpha){
    color <-  list(rgb(0, 255, 170, max = 255, alpha = alpha, names ='green'),
                   rgb(0, 221, 255, max = 255, alpha = alpha, names='sky blue'),
                   rgb(255, 0, 255, max = 255, alpha = alpha, names='purple'),
                   rgb(255, 157, 0, max = 255, alpha = alpha, names='orange'),
                   rgb(255, 0, 212, max = 255, alpha = alpha, names='magenta' ))
    return(color)
 }

 # Assign period to data
 obs_segmented <- data.frame()

 # Conditional to separate non segmentation case
 if(typeof(res$data.p$obs.p)=='list'){
   for(i in 1:length(res$data.p$obs.p)){
    obs_segmented_temp=cbind(obs=res$data.p$obs.p[[i]],period=i)
    obs_segmented=rbind(obs_segmented,obs_segmented_temp)
   }
 }else{
  obs_segmented=data.frame(obs=res$data.p$obs.p,period=1)
 }

 # Plot observations
 plot(x=obs_segmented$obs,
      col=factor(obs_segmented$period),
      pch=16,
      main='Final segmentation',
      ylab='obs',
      xlab='time')

 # Plot segments
 for(i in 1:nS.user){
   segments(x0=intervals.time.shift[[i]],
            x1=intervals.time.shift[[i+1]],
            y0=res$segments[[i]],
            y1=res$segments[[i]],
            col='blue')
   rect(xleft=intervals.time.shift[[i]],
        xright=intervals.time.shift[[i+1]],
        ybottom=uncertainty95_segment[[i]][1],
        ytop=uncertainty95_segment[[i]][2],
        col= rgb(0,0,255,max=255,alpha=125,names='blue'),
        border = 'transparent')
 }

 # Plot shifts
 if(nS.user!=1){
  for(i in 1:(nS.user-1)){
   abline(v=res$tau[i],col=color_customized_rect(255)[[i]], lwd=2)
   rect(xleft=uncertainty95_shift[[i]][1],
        xright=rev(uncertainty95_shift[[i]])[1],
        ybottom=min(obs)*2,
        ytop=max(obs)*2,
        col= color_customized_rect(125)[[i]],
        border = 'transparent')
   }
 }
```

##  Segmentation procedure for a **unknown** number of segments

This is a basic example which shows you how to segment a variable with a **unknown** number of segments :

```{r segmentation}
 # Set random generation
 set.seed(1)

 # Create observation vector
 obs=c(rnorm(25,mean=0,sd=1),rnorm(25,mean=2,sd=1))
 
 # Set the maximum number of segments 
 nSmax.user=3

 # Run segmentation function
 res <- segmentation(obs=obs,nSmax=nSmax.user)

 # Optimal number of segments nSopt
 nSopt <- res$nS
 nSopt

 # Estimated shift time
 res$results[[nSopt]]$tau

 # intervals defined by time shifts
 if(nSopt!=1){
  intervals.time.shift=c(res$results[[nSopt]]$data.p$time[[1]][1],
                         res$results[[nSopt]]$tau,
                         rev(res$results[[nSopt]]$data.p$time[[nSopt]])[1])
 }else{
  intervals.time.shift=list(res$results[[nSopt]]$data.p$time[1],
                            rev(res$results[[nSopt]]$data.p$time)[1])
 }

 # Maximum a posterior value per segment indexed by the list number
 res$results[[nSopt]]$segments

 # Uncertainty in shift time
 if(nSopt!=1){
    Shift=res$results[[nSopt]]$mcmc$tau1
    hist(Shift,
         main='Histogram of first shift')

    uncertainty95_shift <- list()
    for(i in 1:(nSopt-1)){
     uncertainty95_shift[[i]] = stats::quantile(res$results[[nSopt]]$mcmc[,nSopt+i],
                                                probs=c(0.025,0.975))
    }
 }

 # Uncertainty in segment estimation
 mu.seg.1.unc=res$results[[nSopt]]$mcmc$mu1
 graphics::hist(mu.seg.1.unc,
                xlab='obs',
                main='Histogram of first segment of observation')

 uncertainty95_segment <- list()
 for(i in 1:nSopt){
    uncertainty95_segment [[i]] = stats::quantile(res$results[[nSopt]]$mcmc[,i],
                                                  probs=c(0.025,0.975))
 }
 # Separate and assign information by identified stable period
 res$results[[nSopt]]$data.p

 # DIC estimation
 res$results[[nSopt]]$DIC

 # Setting plot

 # Transparency
 alpha <- 125

 # Set color plot
 color_customized_rect <- function(alpha){
    color <-  list(rgb(0, 255, 170, max = 255, alpha = alpha, names ='green'),
                   rgb(0, 221, 255, max = 255, alpha = alpha, names='sky blue'),
                   rgb(255, 0, 255, max = 255, alpha = alpha, names='purple'),
                   rgb(255, 157, 0, max = 255, alpha = alpha, names='orange'),
                   rgb(255, 0, 212, max = 255, alpha = alpha, names='magenta' ))
    return(color)
 }

 # Assign period to data
 obs_segmented <- data.frame()

 # Conditional to separate non segmentation case
 if(typeof(res$results[[nSopt]]$data.p$obs.p)=='list'){
   for(i in 1:length(res$results[[nSopt]]$data.p$obs.p)){
    obs_segmented_temp=cbind(obs=res$results[[nSopt]]$data.p$obs.p[[i]],period=i)
    obs_segmented=rbind(obs_segmented,obs_segmented_temp)
   }
 }else{
  obs_segmented=data.frame(obs=res$results[[nSopt]]$data.p$obs.p,period=1)
 }

 # Plot observations
 plot(x=obs_segmented$obs,
      col=factor(obs_segmented$period),
      pch=16,
      main='Final segmentation',
      ylab='obs',
      xlab='time')

 # Plot segments
 for(i in 1:nSopt){
   segments(x0=intervals.time.shift[[i]],
            x1=intervals.time.shift[[i+1]],
            y0=res$results[[nSopt]]$segments[[i]],
            y1=res$results[[nSopt]]$segments[[i]],
            col='blue')
   rect(xleft=intervals.time.shift[[i]],
        xright=intervals.time.shift[[i+1]],
        ybottom=uncertainty95_segment[[i]][1],
        ytop=uncertainty95_segment[[i]][2],
        col= rgb(0,0,255,max=255,alpha=125,names='blue'),
        border = 'transparent')
 }

 # Plot shifts
 if(nSopt!=1){
  for(i in 1:(nSopt-1)){
   abline(v=res$results[[nSopt]]$tau[i],col=color_customized_rect(255)[[i]], lwd=2)
   rect(xleft=uncertainty95_shift[[i]][1],
        xright=rev(uncertainty95_shift[[i]])[1],
        ybottom=min(obs)*2,
        ytop=max(obs)*2,
        col= color_customized_rect(125)[[i]],
        border = 'transparent')
   }
 }
``` 

## recursive.segmentation function

Recursive segmentation procedure for a **unknown** given number of segments

This is a basic example which shows you how to segment a variable with a **unknown** number of segments using a recursive process:

```{r}
# Set random generation
set.seed(1)
# Create series to be segmented
# obs=c(rnorm(30,mean=0,sd=1),rnorm(30,mean=2,sd=1))
# obs=c(rnorm(30,mean=0,sd=1),
#       rnorm(30,mean=2,sd=1),
#       rnorm(30,mean=1,sd=1),
#       rnorm(30,mean=2,sd=1))

obs=c(rnorm(30,mean=0.5,sd=1),
      rnorm(30,mean=1.5,sd=1),
      rnorm(30,mean=1,sd=1),
      rnorm(30,mean=2,sd=1))

time=1:length(obs)

# Assign a maximum number of user-defined segments to be assessed
nSmax.user=3

# Apply recursive segmentation
results=recursive.segmentation(obs, nSmax=nSmax.user)

# Have a look at recursion tree
results$tree

# Visualize tree with data.tree package
# if(NROW(results$tree)>1){
#   tree <- data.tree::as.Node(results$tree[-1,c(3,1)],mode = "network")
#   plot(tree)
# } else { # No segmentation took place, make a dummy plot
#   tree <- data.tree::as.Node(data.frame(1,2),mode = "network")
#   plot(tree)
# }

# Get terminal nodes
terminal=which(results$tree$nS==1)

# Get node with time shifts
nodes.shift.time=which(results$tree$nS!=1)

nodes.shift.time

# Estimated shift time along with uncertainties
shift.time.list <- c()
for(i in 1:length(nodes.shift.time)){
  nSopt.p = results$res[[nodes.shift.time[[i]]]]$nS
  results.p=results$res[[nodes.shift.time[[i]]]]$results
  # shift.time.p=results.p[[nSopt.p]]$tau
  shift.time.p=cbind(c(results.p[[nSopt.p]]$tau))

  for(j in 1:(nSopt.p-1)){

    shift.time.p.unc=data.frame(tau=as.numeric(shift.time.p[j,]),
                                u2.5=stats::quantile(results.p[[nSopt.p]]$mcmc[,nSopt.p+j],
                                                     probs=c(0.025)),
                                u97.5=stats::quantile(results.p[[nSopt.p]]$mcmc[,nSopt.p+j],
                                                      probs=c(0.975)))
    shift.time.list <- rbind(shift.time.list,
                         shift.time.p.unc)
  }
}

all.shift.time <- shift.time.list[order(shift.time.list$tau),]
all.shift.time

# intervals defined by time shifts
intervals.time.shift=c(results$res[[1]]$results[[1]]$data.p$time.p[1],
                       all.shift.time$tau,
                       rev(results$res[[1]]$results[[1]]$data.p$time.p)[1])

# Get stable periods by adding information as period, segment,
data.stable <- c()
for(i in 1:length(terminal)){
  data.stable.p=results$res[[terminal[i]]]$results[[1]]   #Save data from stable period

  node = data.frame(obs=data.stable.p$data.p$obs.p,
                    time=data.stable.p$data.p$time.p,
                    u=data.stable.p$data.p$u.p,
                    period = rep(i,length(data.stable.p$data.p$obs.p)))
  data.stable = rbind(data.stable,node)
}

# Setting plot

# Transparency
alpha <- 125

# Set color plot
color_customized_rect <- function(alpha){
  color <-  list(rgb(0, 255, 170, max = 255, alpha = alpha, names ='green'),
                 rgb(0, 221, 255, max = 255, alpha = alpha, names='sky blue'),
                 rgb(255, 0, 255, max = 255, alpha = alpha, names='purple'),
                 rgb(255, 157, 0, max = 255, alpha = alpha, names='orange'),
                 rgb(255, 0, 212, max = 255, alpha = alpha, names='magenta' ))
  return(color)
}

# Plot observations
plot(x=data.stable$time,
y=data.stable$obs,
col=data.stable$period,
xlab='time',
ylab='obs',
main='Final segmentation'
)

# Plot shifts
for(i in 1:nrow(all.shift.time)){
  abline(v=all.shift.time$tau[i],
         col=color_customized_rect(255)[[i]])
  rect(xleft=all.shift.time$u2.5[[i]],
       xright=all.shift.time$u97.5[[i]],
       ybottom=min(obs)*2,
       ytop=max(obs)*2,
       col= color_customized_rect(50)[[i]],
       border = 'transparent')
}

# Pre-treatment data to extract segments from corresponding node
all.global.data <- c()
for(i in 1:length(nodes.shift.time)){
  nSopt.p = results$res[[nodes.shift.time[[i]]]]$nS
  results.p=results$res[[nodes.shift.time[[i]]]]$results

  relation.parent.children.p=results$tree[which(results$tree$parent==nodes.shift.time[[i]]),]
  id.segment.p=which(relation.parent.children.p$nS==1)

  for(j in 1:length(id.segment.p)){
    segments.p=results.p[[nSopt.p]]$segments[[id.segment.p[[j]]]]
    obs.p=results.p[[nSopt.p]]$data.p$obs.p[[id.segment.p[[j]]]]
    time.p=results.p[[nSopt.p]]$data.p$time.p[[id.segment.p[[j]]]]
    u.p=results.p[[nSopt.p]]$data.p$u.p[[id.segment.p[[j]]]]
    unc.segment.p=stats::quantile(results.p[[nSopt.p]]$mcmc[,id.segment.p[[j]]],
                                  probs=c(0.025,0.975))

    data.p.temp=data.frame(mu=segments.p,
                           obs=obs.p,
                           u=u.p,
                           time=time.p,
                           u2.5=rep(unc.segment.p[1],length(segments.p)),
                           u97.5=rep(unc.segment.p[2],length(segments.p)))

    all.global.data=rbind(all.global.data,data.p.temp)
  }
}

all.global.data=all.global.data[order(all.global.data$time),]

final.data.segmented=unname((split(all.global.data,all.global.data$mu)))

# Plot segments
for(i in 1:(nrow(all.shift.time)+1)){
  text(final.data.segmented[[i]]$time,
       final.data.segmented[[i]]$obs,
       terminal[i],pos=3,
       cex=0.8,
       col='black')
  segments(x0=intervals.time.shift[[i]],
           x1=intervals.time.shift[[i+1]],
           y0=final.data.segmented[[i]]$mu,
           y1=final.data.segmented[[i]]$mu,
           col='blue')
  rect(xleft=intervals.time.shift[[i]],
       xright=intervals.time.shift[[i+1]],
       ybottom=final.data.segmented[[i]]$u2.5,
       ytop=final.data.segmented[[i]]$u97.5,
       col= rgb(0,0,255,max=255,alpha=5,names='blue'),
       border = 'transparent')

}
```


