---
title: "Rating Shift Happens"
author: "Felipe MENDEZ and Benjamin RENARD (INRAE, RiverLy and RECOVER)."
date: "December 2023"
output:
  github_document:
subtitle: Documentation
bibliography: bibliography/RatingShiftHappens.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  comment = "#>",
  fig.path = "man/readme/README-",
  out.width = "100%"
)
```

## Introduction

The goal of `RatingShiftHappens` package is to create a tools package for detecting, visualizing and estimating rating shifts. This package was derived from [BayDERS](https://github.com/MatteoDarienzo/BayDERS) developed by [@darienzoDetectionEstimationStagedischarge2021](https://theses.hal.science/tel-03211343).

This documentation provides the description of several functions available to the segmentation process. 

Three fundamental functions are available to segment a random variable :

1. segmentation.engine
2. segmentation
3. recursive.segmentation

## Installation

You can install the development version of [RatingShiftHappens](https://github.com/Felipemendezrios/RatingShiftHappens) from GitHub using the following command. Please note that the [RBaM](https://github.com/BaM-tools/RBaM) package developed by [@renardBaMBayesianModeling2017](https://hal.inrae.fr/hal-02606929) is also required to run this package.

``` {r  results='hide',message=FALSE}
# Before first use, install Rating Shift Happens and RBaM packages ones and for all, following these commands: 

# devtools::install_github("Felipemendezrios/RatingShiftHappens")
# devtools::install_github('BaM-tools/RBaM') 

library(RatingShiftHappens)
```

Functions will be explained more precisely below along with an example.

##  Segmentation procedure for a *known* given number of segments

This basic example demonstrates the segmentation of annual maximum stages (H, m) for the Rhone River at Beaucaire, France, along with the associated uncertainties expressed as standard deviations (uH), divided into two groups. More information about the data set, please refer to the documentation available in `?RhoneRiver`.

```{r segmentation.engine}
 # Run segmentation engine function at two segments
 res=segmentation.engine(obs=RhoneRiver$H,
                         time=RhoneRiver$Year,
                         u=RhoneRiver$uH,
                         nS=2)
 # Data information
 knitr::kable(head(res$summary$data),
              align = 'c',row.names = F)
 # Shift information
 knitr::kable(head(res$summary$shift),
              align = 'c',row.names = F)
 # Plot segmentation
 Plots=plotSegmentation(summary=res$summary,
                        plot_summary=res$plot)
 
 # Observations of the random variable and shift time estimated 
 Plots$observation_and_shift
 
 # Probability distribution function for detecting shift time with a 95% credibility interval 
 Plots$shift_time_density
 
 # Final plot segmentation
 Plots$final_plot
```



## For more advanced details : 

MCMC sampling demonstrate all combinations of parameters estimated. 

```{r}
knitr::kable(head(res$mcmc),align = 'c')

```

The `RBaM` package provides several functions to explore MCMC samples.

```{r}
  # Trace plot for each parameter, useful to assess convergence.
  plots=RBaM::tracePlot(res$mcmc)
  gridExtra::grid.arrange(grobs=plots,ncol=3)
  # Density plot for each parameter
  plots=RBaM::densityPlot(res$mcmc)
  gridExtra::grid.arrange(grobs=plots,ncol=3)

```


In this example, the focus will be on exploring the uncertainty associated with the first shift time. 

```{r}
 Shift=data.frame(time=res$mcmc$tau1)
 ggplot2::ggplot(Shift,ggplot2::aes(x=time))+
   ggplot2::geom_histogram(ggplot2::aes(y=..density..),col=1,fill='white',bins=80)+
   ggplot2::labs(title='Histogram with density of first shift')+
   ggplot2::theme_bw()+
   ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))+
   ggplot2::geom_density(col=4,lwd=1,fill=4,alpha=0.25)
```
 
##  Segmentation procedure for an *unknown* number of segments

This is a basic example, which shows you how to segment the same dataset with an **unknown** number of segments :

```{r segmentation}
 # Run segmentation engine function at two segments
 res=segmentation(obs=RhoneRiver$H,
                  time=RhoneRiver$Year,
                  u=RhoneRiver$uH,
                  nSmax=3)

 # Get lower DIC value and optimal number of segments (to define optimal solution)
 DIC.df = data.frame(nS=c(1:3),DIC=c(res$results[[1]]$DIC,res$results[[2]]$DIC,res$results[[3]]$DIC))
 nSopt=res$nS
 
 ggplot2::ggplot(DIC.df,ggplot2::aes(x=nS,y=DIC,col=factor(nS)))+
   ggplot2::geom_point(size=3,show.legend = F)+
   ggplot2::annotate('segment',
                     x=nSopt,y=min(DIC.df$DIC)*1.03,xend=nSopt,yend=min(DIC.df$DIC)*1.01,
                     linewidth=2,linejoin = "mitre",
                     arrow=ggplot2::arrow(type='closed',length=ggplot2::unit(0.01,'npc')))+
   ggplot2::theme_bw()
 # Data information
 knitr::kable(head(res$results[[nSopt]]$summary$data),
              align = 'c',row.names = F)
 # Shift information
 knitr::kable(head(res$results[[nSopt]]$summary$shift),
              align = 'c',row.names = F)
 # Final plot segmentation
 plotSegmentation(summary=res$summary,
                  plot_summary = res$plot)$final_plot
``` 

## Recursive segmentation procedure for an *unknown*  number of segments

This is a basic example, which shows you how to segment the data set with an **unknown** number of segments using a recursive process:

```{r recursive.segmentation}
 # Apply recursive segmentation
 results=recursive.segmentation(obs=RhoneRiver$H,
                                time=RhoneRiver$Year,
                                u=RhoneRiver$uH,
                                nSmax=3)
 
 # Data information
 knitr::kable(head(results$summary$data),
              align = 'c',row.names = F)
 # Shift information
 knitr::kable(head(results$summary$shift),
              align = 'c',row.names = F)
 # Have a look at recursion tree
 results$tree

 # Visualize tree structure
 plotTree(tree=results$tree)
 
 # Final plot segmentation
 plotSegmentation(summary=results$summary,
                  plot_summary = results$plot)$final_plot
```

## Hydrometry field

Detection and segmentation has only been performed for the residual of a random variable thus far. However, in the field of hydrometry,one of the objectives is to predict discharge from stage, using a rating curve that can vary over time.

## Fitting models

Many models are available to describe the rating curve. All fitting models with their equations supported by the package are listed below. For more details, refer to `GetCatalog()` to determine which model could be used to estimate the rating curve. 

```{r}
# Get model available to estimate the rating curve
GetCatalog()$models

# Get equation of each model
GetCatalog()$Equations
```
All these equations Q(h) allow for the proper transformation of stage to discharge, following the specified assumption for each fitting model.

Models can either be non-parametric, such as as `fitRC_loess`, which relies solely on data for calculation, or parametric, like `fitRC_BaRatin` with three parameters (a,b,c) per hydraulic control, integrating physics and geometry proprieties of the river in the estimation process. 

Hereafter, the employed model will be an exponential regression (`fitRC_exponential`) and the BaRatin model for estimating discharge.

The exponential regression needs two parameters, denoted as *a* and *b*, following the equation :

$Q(h) = a \cdot e^{(b \cdot h)}$

The BaRatin model needs the parameters *a*, *b* and *c* per hydraulic control, following the equation : 

$Q(h) = a \cdot (h-b)^{c} \quad \text{for } (h>k) \quad (\text{and } Q=0 \quad \text{if } h \leq b)$ 

## Dataset

The Ardèche hydrometric station at Meyras is introduced as a new dataset, further information in `?ArdecheRiverMeyrasGaugings`. The dataset includes stages (H, in meters) and discharge ADCP measurements (Q, in cubic meters per second) all accompanied by uncertainties. 

```{r}
knitr::kable(head(ArdecheRiverMeyrasGaugings),
              align = 'c',row.names = F)
```

## Recursive model and segmentation procedure for an *unknown*  number of segments

This function enables the modeling of the rating curve and ensures its continual update at each segmentation for an **unknown** number of segments. This approach leads to a better fit for the model as it is consistently updated with data from the current period.

### Rating curve using exponential regression

An exponential regression model is employed  to construct the rating curve using observed data point represented by stage and discharge information. This regression estimates the relationships between a dependent variable and one independent variables from a statistical perspective.

```{r recursive.ModelAndSegmentation}
# Apply recursive model and segmentation
results=recursive.ModelAndSegmentation(H=ArdecheRiverMeyrasGaugings$H,Q=ArdecheRiverMeyrasGaugings$Q,
                                       time=ArdecheRiverMeyrasGaugings$Date,
                                       uQ=ArdecheRiverMeyrasGaugings$uQ,
                                       nSmax=2,nMin=2,funk=fitRC_exponential)

# Data information
knitr::kable(head(results$summary$data),
             align = 'c',row.names = FALSE)

# Shift information
knitr::kable(head(results$summary$shift),
            align = 'c',row.names = FALSE)

# Parameters estimation of the rating curve
results$summary$param.equation

# Have a look at recursion tree
results$tree

```

Several plot functions are available to simplify graphical representation, aligning with the structure of specific functions integrated in the package: 

```{r}
# Visualize tree structure
plotTree(results$tree)

# parameter of the rating curve 
a=results$summary$param.equation$a
b=results$summary$param.equation$b

# Plot the rating curve after segmentation following a regression exponential
plotRC_ModelAndSegmentation(summary=results$summary,
                            equation = Exponential_Equation,
                            a=a,
                            b=b)

# Plot the rating curves after segmentation with zoom user-defined
plotRC_ModelAndSegmentation(summary=results$summary,
                            equation = Exponential_Equation, 
                            a=a,
                            b=b,
                            autoscale = FALSE, 
                            Hmin_user = 1, 
                            Hmax_user = 2,
                            H_step_discretization = 0.01)

# Plot the rating curves after segmentation in log scale
plotRC_ModelAndSegmentation(summary=results$summary,
                            logscale=TRUE,
                            equation = Exponential_Equation,
                            a=a,
                            b=b)

# Plot the rating curves after segmentation in log scale with zoom
plotRC_ModelAndSegmentation(summary=results$summary,
                            logscale=TRUE,
                            equation = Exponential_Equation,
                            a=a,
                            b=b,
                            autoscale = FALSE, 
                            Hmin_user = 0.5, 
                            Hmax_user = 2, 
                            H_step_discretization = 0.01)

# Plot shift times in stage record
plot_H_ModelAndSegmentation(summary=results$summary,
                            plot_summary=results$plot)$final_plot

# Plot shift times in discharge observations
plot_Q_ModelAndSegmentation(summary=results$summary,
                            plot_summary=results$plot)$final_plot

# Plot residual
plotResidual_ModelAndSegmentation(summary=results$summary,
                                  plot_summary=results$plot)$final_plot

```

### Rating curve using BaRatin method

The Bayesian BaRatin method ([Bayesian Rating curve](https://baratin-tools.github.io/en/), @lecozCombiningHydraulicKnowledge2014; @hornerImpactStageMeasurement2018) has been developed at INRAE. 

The method combines the strength of a probabilistic approach (parameter estimation and uncertainty quantification) and a physically-based approach (physical interpretation of parameters and of their change when a rating shift occurs, more reliable extrapolation). BaRatin provides a way for field hydrologists to simply formalize and make use of the expertise they do have on flows at their gauging stations, along with gaugings and their uncertainties.

In this package, the methodology of BaRatin will not been explained in detail. Instead, hydraulic control matrix and prior information on parameters will be used to run the segmentation and estimation of the rating curve.

#### Hydraulic analysis

Firstly, a hydraulic analysis of the gauging station is required to set the hydraulic control matrix. The Ardèche River at Meyras station is of interest because it illustrates a frequently-encountered 3-control configuration (riffle, main channel, floodway).

At low flows, the stage-discharge relation is controlled by the geometry of a critical section induced by a natural riffle.

As stage increases, the riffle becomes drowned and the stage-discharge relation is controlled by the geometry and roughness of the main channel. 

At high flows, part of the water flows into two floodways located on the right and left banks. Since the two floodways get activated at roughly the same stage, they are combined into a single control. 

To assist the user in entering the hydraulic control matrix, the `control_matrix_builder` function was developed. This function interacts with the user to facilitate the creation of the matrix.

```{r}
# Hydraulic control matrix
controlMatrix=matrix(c(1,0,0,0,1,1,0,0,1),ncol=3,nrow=3)
```

#### Prior information 

The method required prior information on the parameters *a*, *b* and *c* per hydraulic control following this equation:


$Q(h) = a*(h-b)^{c} \quad \text{for } (h>k) \quad (\text{and } Q=0 \quad \text{if } h \leq b)$  

* Parameter *a* is the coefficient representing the geometry and physical properties of the control. It will be estimated differently in function of the type of control. 

* Parameter *b* is the offset; when stage falls below the value *b*, discharge id zero.

* Parameter *c* is the exponent, which depends solely on the type of control. 

* Parameter *k* is the activation stage; when the water level falls below the value *k*, the control becomes inactive.

See the details of the values entered here : [prior specification for the case of study of Ardeche at Meyras gauging station](https://baratin-tools.github.io/en/doc/case/ardeche-meyras/#prior-specification).

```{r}
# Set prior information to each hydraulic control 
a1=RBaM::parameter(name='a1',init=14.17,prior.dist='LogNormal',prior.par=c(2.66,1.54))
b1=RBaM::parameter(name='b1',init=-0.6,prior.dist='Gaussian',prior.par=c(-0.58,1.49))
c1=RBaM::parameter(name='c1',init=1.5,prior.dist='Gaussian',prior.par=c(1.5,0.025))
a2=RBaM::parameter(name='a2',init=26.5165,prior.dist='LogNormal',prior.par=c(3.28,0.36))
b2=RBaM::parameter(name='b2',init=-0.6,prior.dist='Gaussian',prior.par=c(-0.58,1.49))
c2=RBaM::parameter(name='c2',init=1.67,prior.dist='Gaussian',prior.par=c(1.67,0.025))
a3=RBaM::parameter(name='a3',init=31.82,prior.dist='LogNormal',prior.par=c(3.46,0.397))
b3=RBaM::parameter(name='b3',init=1.2,prior.dist='Gaussian',prior.par=c(1.2,0.2))
c3=RBaM::parameter(name='c3',init=1.67,prior.dist='Gaussian',prior.par=c(1.67,0.025))

# Set a list of the same parameters for all controls
a.object=list(a1,a2,a3)
b.object=list(b1,b2,b3)
c.object=list(c1,c2,c3)
```

The `prior_infor_param_builder` function was developed and available to help the user to create these objects in a interactive way.

```{r}
# Apply recursive model and segmentation with BaRatin multi-control method
resultsBaRatin=recursive.ModelAndSegmentation(H=ArdecheRiverMeyrasGaugings$H,
                                              Q=ArdecheRiverMeyrasGaugings$Q,
                                              time=ArdecheRiverMeyrasGaugings$Date,
                                              uQ=ArdecheRiverMeyrasGaugings$uQ,
                                              nSmax=3,
                                              nMin=2,
                                              funk=fitRC_BaRatin,
                                              HmaxGrid=max(ArdecheRiverMeyrasGaugings$H),
                                              a.object=a.object,
                                              b.object=b.object,
                                              c.object=c.object,
                                              controlMatrix=controlMatrix
                                              )

 # Visualize tree structure
 plotTree(resultsBaRatin$tree)

 # Terminal nodes
 terminal = resultsBaRatin$tree$indx[which(resultsBaRatin$tree$nS==1)]
 terminal
```

Plot the rating curves after using BaRatin method. It is possible to plot all nodes in the tree structure by setting `allnodes=TRUE`. To reduce calculation time, it is advisable to specify the vector of nodes for plotting the rating curve.

```{r}
PlotRCPrediction(Hgrid=data.frame(seq(-1,2,by=0.01)),
                  autoscale=FALSE,
                  temp.folder=file.path(tempdir(),'BaM'),
                  CalibrationData='CalibrationData.txt',
                  allnodes=FALSE,
                  nodes=terminal)

 # Plot shift times in stage record
 plot_H_ModelAndSegmentation(summary=resultsBaRatin$summary,
                             plot_summary=resultsBaRatin$plot)$final_plot

 # Plot shift times in discharge observations
 plot_Q_ModelAndSegmentation(summary=resultsBaRatin$summary,
                             plot_summary=resultsBaRatin$plot)$final_plot

 # Plot residual
 plotResidual_ModelAndSegmentation(summary=resultsBaRatin$summary,
                                   plot_summary=resultsBaRatin$plot)$final_plot
```
 
## Références

