---
title: "Rating Shift Happens"
subtitle: "Documentation"
author: "Felipe MENDEZ and Benjamin RENARD (INRAE, RiverLy and RECOVER)."
date: "December 2023"
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/readme/README-",
  out.width = "100%"
)
```

## Introduction

Darienzo (2021) developed [BayDERS](https://github.com/MatteoDarienzo/BayDERS) for detecting and estimating stage-discharge rating shifts for retrospective and real-time streamflow quantification. In 2022, a sensibility analysis was conducted to assess the stability and performance of the tool, leading to multiple corrections. Nevertheless, debugging the tool proved difficult due to the integration of different models such as rating curve and recessions estimation. Hence, a dissociation procedure has been implemented to extract essential functions, enabling the automatic segmentation of a random variable. Theses functions have been organized and coded in a package format, which will be explained below.   

The goal of `RatingShiftHappens` package is to create a tools package for detecting, visualizing and estimating rating shifts. This documentation provides the description of several functions available to the segmentation process. 

Three fundamental functions are available so far :


1. segmentation.engine
2. segmentation
3. recursive.segmentation

## Installation

You can install the development version of [RatingShiftHappens](https://github.com/Felipemendezrios/RatingShiftHappens) from GitHub using the following command. Please note that the [RBaM](https://github.com/BaM-tools/RBaM) package is also required to run this package.

``` {r  results='hide',message=FALSE}
# install.packages("devtools")
devtools::install_github("Felipemendezrios/RatingShiftHappens")
devtools::install_github('BaM-tools/RBaM') 

library(RatingShiftHappens)
```

Functions will be explained more precisely below along with a example.

##  segmentation.engine function 

Segmentation procedure for a **known** given number of segments

This is a basic example which shows you how to segment a random variable:

```{r segmentation.engine}
 # Set random generation
 set.seed(1)

 # Create observation vector
 obs=c(rnorm(25,mean=0,sd=1),rnorm(25,mean=2,sd=1))
 nS.user=2

 # Run segmentation engine function
 res <- segmentation.engine(obs=obs,nS=nS.user)

 # Estimated shift time
 res$tau

 # intervals defined by time shifts
 if(nS.user!=1){
  intervals.time.shift=c(res$data.p$time[[1]][1],res$tau,rev(res$data.p$time[[nS.user]])[1])
 }else{
  intervals.time.shift=list(res$data.p$time[1],rev(res$data.p$time)[1])
 }

 # Maximum a posterior value per segment indexed by the list number
 res$segments

 # Uncertainty in shift time
 if(nS.user!=1){
    Shift=res$mcmc$tau1
    hist(Shift,
         main='Histogram of first shift')

    uncertainty95_shift <- list()
    for(i in 1:(nS.user-1)){
     uncertainty95_shift[[i]] = stats::quantile(res$mcmc[,nS.user+i],probs=c(0.025,0.975))
    }
 }


 # Uncertainty in segment estimation
 mu.seg.1.unc=res$mcmc$mu1
 graphics::hist(mu.seg.1.unc,
                xlab='obs',
                main='Histogram of first segment of observation')

 uncertainty95_segment <- list()
 for(i in 1:nS.user){
    uncertainty95_segment [[i]] = stats::quantile(res$mcmc[,i],probs=c(0.025,0.975))
 }

 # Separate and assign information by identified stable period
 res$data.p
 
 # DIC estimation
 res$DIC

 # Setting transparency for plotting
 alpha = 125

 # Set color plot
 color_customized_rect <- function(alpha){
    color <-  list(rgb(0, 255, 170, max = 255, alpha = alpha, names ='green'),
                   rgb(0, 221, 255, max = 255, alpha = alpha, names='sky blue'),
                   rgb(255, 0, 255, max = 255, alpha = alpha, names='purple'),
                   rgb(255, 157, 0, max = 255, alpha = alpha, names='orange'),
                   rgb(255, 0, 212, max = 255, alpha = alpha, names='magenta' ))
    return(color)
 }

 # Assign period to data
 obs_segmented <- data.frame()

 # Conditional to separate non segmentation case
 if(typeof(res$data.p$obs.p)=='list'){
   for(i in 1:length(res$data.p$obs.p)){
    obs_segmented_temp=cbind(obs=res$data.p$obs.p[[i]],period=i)
    obs_segmented=rbind(obs_segmented,obs_segmented_temp)
   }
 }else{
  obs_segmented=data.frame(obs=res$data.p$obs.p,period=1)
 }

 # Plot observations
 plot(x=obs_segmented$obs,
      col=factor(obs_segmented$period),
      pch=16,
      main='Final segmentation',
      ylab='obs',
      xlab='time')

 # Plot segments
 for(i in 1:nS.user){
   segments(x0=intervals.time.shift[[i]],
            x1=intervals.time.shift[[i+1]],
            y0=res$segments[[i]],
            y1=res$segments[[i]],
            col='blue')
   rect(xleft=intervals.time.shift[[i]],
        xright=intervals.time.shift[[i+1]],
        ybottom=uncertainty95_segment[[i]][1],
        ytop=uncertainty95_segment[[i]][2],
        col= rgb(0,0,255,max=255,alpha=125,names='blue'),
        border = 'transparent')
 }

 # Plot shifts
 if(nS.user!=1){
  for(i in 1:(nS.user-1)){
   abline(v=res$tau[i],col=color_customized_rect(255)[[i]], lwd=2)
   rect(xleft=uncertainty95_shift[[i]][1],
        xright=rev(uncertainty95_shift[[i]])[1],
        ybottom=min(obs)*2,
        ytop=max(obs)*2,
        col= color_customized_rect(125)[[i]],
        border = 'transparent')
   }
 }
```

##  segmentation function 

Segmentation procedure for a **unknown** given number of segments

This is a basic example which shows you how to segment a variable with a **unknown** number of segments:

```{r segmentation}
 # Set random generation
 set.seed(1)

 # Create observation vector
 obs=c(rnorm(25,mean=0,sd=1),rnorm(25,mean=2,sd=1))
 
 # Set the maximum number of segments 
 nSmax.user=3

 # Run segmentation function
 res <- segmentation(obs=obs,nSmax=nSmax.user)

 # Optimal number of segments nSopt
 nSopt <- res$nS
 nSopt

 # Estimated shift time
 res$results[[nSopt]]$tau

 # intervals defined by time shifts
 if(nSopt!=1){
  intervals.time.shift=c(res$results[[nSopt]]$data.p$time[[1]][1],
                         res$results[[nSopt]]$tau,
                         rev(res$results[[nSopt]]$data.p$time[[nSopt]])[1])
 }else{
  intervals.time.shift=list(res$results[[nSopt]]$data.p$time[1],
                            rev(res$results[[nSopt]]$data.p$time)[1])
 }

 # Maximum a posterior value per segment indexed by the list number
 res$results[[nSopt]]$segments

 # Uncertainty in shift time
 if(nSopt!=1){
    Shift=res$results[[nSopt]]$mcmc$tau1
    hist(Shift,
         main='Histogram of first shift')

    uncertainty95_shift <- list()
    for(i in 1:(nSopt-1)){
     uncertainty95_shift[[i]] = stats::quantile(res$results[[nSopt]]$mcmc[,nSopt+i],
                                                probs=c(0.025,0.975))
    }
 }

 # Uncertainty in segment estimation
 mu.seg.1.unc=res$results[[nSopt]]$mcmc$mu1
 graphics::hist(mu.seg.1.unc,
                xlab='obs',
                main='Histogram of first segment of observation')

 uncertainty95_segment <- list()
 for(i in 1:nSopt){
    uncertainty95_segment [[i]] = stats::quantile(res$results[[nSopt]]$mcmc[,i],
                                                  probs=c(0.025,0.975))
 }
 # Separate and assign information by identified stable period
 res$results[[nSopt]]$data.p

 # DIC estimation
 res$results[[nSopt]]$DIC

 # Setting plot

 # Transparency
 alpha <- 125

 # Set color plot
 color_customized_rect <- function(alpha){
    color <-  list(rgb(0, 255, 170, max = 255, alpha = alpha, names ='green'),
                   rgb(0, 221, 255, max = 255, alpha = alpha, names='sky blue'),
                   rgb(255, 0, 255, max = 255, alpha = alpha, names='purple'),
                   rgb(255, 157, 0, max = 255, alpha = alpha, names='orange'),
                   rgb(255, 0, 212, max = 255, alpha = alpha, names='magenta' ))
    return(color)
 }

 # Assign period to data
 obs_segmented <- data.frame()

 # Conditional to separate non segmentation case
 if(typeof(res$results[[nSopt]]$data.p$obs.p)=='list'){
   for(i in 1:length(res$results[[nSopt]]$data.p$obs.p)){
    obs_segmented_temp=cbind(obs=res$results[[nSopt]]$data.p$obs.p[[i]],period=i)
    obs_segmented=rbind(obs_segmented,obs_segmented_temp)
   }
 }else{
  obs_segmented=data.frame(obs=res$results[[nSopt]]$data.p$obs.p,period=1)
 }

 # Plot observations
 plot(x=obs_segmented$obs,
      col=factor(obs_segmented$period),
      pch=16,
      main='Final segmentation',
      ylab='obs',
      xlab='time')

 # Plot segments
 for(i in 1:nSopt){
   segments(x0=intervals.time.shift[[i]],
            x1=intervals.time.shift[[i+1]],
            y0=res$results[[nSopt]]$segments[[i]],
            y1=res$results[[nSopt]]$segments[[i]],
            col='blue')
   rect(xleft=intervals.time.shift[[i]],
        xright=intervals.time.shift[[i+1]],
        ybottom=uncertainty95_segment[[i]][1],
        ytop=uncertainty95_segment[[i]][2],
        col= rgb(0,0,255,max=255,alpha=125,names='blue'),
        border = 'transparent')
 }

 # Plot shifts
 if(nSopt!=1){
  for(i in 1:(nSopt-1)){
   abline(v=res$results[[nSopt]]$tau[i],col=color_customized_rect(255)[[i]], lwd=2)
   rect(xleft=uncertainty95_shift[[i]][1],
        xright=rev(uncertainty95_shift[[i]])[1],
        ybottom=min(obs)*2,
        ytop=max(obs)*2,
        col= color_customized_rect(125)[[i]],
        border = 'transparent')
   }
 }
``` 

## recursive.segmentation function

Recursive segmentation procedure for a **unknown** given number of segments

This is a basic example which shows you how to segment a variable with a **unknown** number of segments using a recursive process:
<!-- #  -->
<!-- #  -->
<!-- # ```{r recursive.segmentation} -->
<!-- # # Set random generation -->
<!-- # set.seed(1) -->
<!-- #  -->
<!-- # # Create series to be segmented -->
<!-- # obs=c(rnorm(30,mean=0,sd=1),rnorm(30,mean=2,sd=1)) -->
<!-- # time=1:length(obs) -->
<!-- #  -->
<!-- # # Apply recursive segmentation -->
<!-- # results=recursive.segmentation(obs) -->
<!-- # # Have a look at recursion tree -->
<!-- # results$tree -->
<!-- #  -->
<!-- # # Get terminal nodes -->
<!-- # terminal=which(results$tree$nS==1) -->
<!-- #  -->
<!-- # # Plot original series and terminal nodes defining final segments -->
<!-- # X11();plot(time,obs) -->
<!-- # for(i in 1:length(terminal)){ -->
<!-- #   data.stable.p=results$res[[terminal[i]]]$results[[1]]   #Save data from stable period -->
<!-- #   node=list(obs=data.stable.p$data.p$obs.p, -->
<!-- #             times=data.stable.p$data.p$time.p, -->
<!-- #             u=data.stable.p$data.p$u.p) -->
<!-- #   points(node$times,node$obs,col=i) -->
<!-- #   text(node$times,node$obs,terminal[i],pos=3,col=i) -->
<!-- # } -->
<!-- # # Get time shifts -->
<!-- # time.shifts=which(results$tree$nS!=1) -->
<!-- # for(i in 1:length(time.shifts)){ -->
<!-- #   nSopt.p = results$res[[time.shifts[[i]]]]$nS -->
<!-- #   abline(v=results$res[[time.shifts[[i]]]]$results[[nSopt.p]]$tau,col='green') -->
<!-- #   abline(v=quantile(results$res[[time.shifts[[i]]]]$results[[nSopt.p]]$mcmc$tau1,probs=c(0.025,0.975)),col='green',lty=2) -->
<!-- #   segments( -->
<!-- #     x0=results$res[[time.shifts[[i]]]]$results[[nSopt.p]]$data.p$time.p[[1]][1], -->
<!-- #     x1=results$res[[time.shifts[[i]]]]$results[[nSopt.p]]$tau, -->
<!-- #     y0=results$res[[time.shifts[[i]]]]$results[[nSopt.p]]$segments[[1]], -->
<!-- #     y1=results$res[[time.shifts[[i]]]]$results[[nSopt.p]]$segments[[1]], -->
<!-- #     col='blue') -->
<!-- #   segments( -->
<!-- #     x0=results$res[[time.shifts[[i]]]]$results[[nSopt.p]]$tau, -->
<!-- #     x1=rev(results$res[[time.shifts[[i]]]]$results[[nSopt.p]]$data.p$time.p[[2]])[1], -->
<!-- #     y0=results$res[[time.shifts[[i]]]]$results[[nSopt.p]]$segments[[2]], -->
<!-- #     y1=results$res[[time.shifts[[i]]]]$results[[nSopt.p]]$segments[[2]], -->
<!-- #     col='blue') -->
<!-- # } -->
<!-- #  -->
<!-- # # Visualize tree with data.tree package -->
<!-- # if(NROW(results$tree)>1){ -->
<!-- #   tree <- data.tree::as.Node(results$tree[-1,c(3,1)],mode = "network") -->
<!-- #   plot(tree) -->
<!-- # } else { # No segmentation took place, make a dummy plot -->
<!-- #   tree <- data.tree::as.Node(data.frame(1,2),mode = "network") -->
<!-- #   plot(tree) -->
<!-- # } -->
<!-- # ``` -->

Here other example to show _____

<!-- # ```{r example} -->
<!-- #  -->
<!-- # ``` -->
devtools::build_readme()

